\documentclass[10pt]{article}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb}
\usepackage{mathtools}
\usepackage[expansion=false]{microtype}
\usepackage[margin=0.9in]{geometry}
\usepackage{booktabs}
\usepackage[dvipsnames]{xcolor}
\usepackage[colorlinks=true,linkcolor=RoyalBlue,urlcolor=RoyalBlue]{hyperref}
\usepackage{titlesec}
\titleformat{\paragraph}[runin]{\normalfont\bfseries\boldmath}{}{0pt}{}[\,---]
\newcommand{\LU}{\mathbf{L}_{U}}
\setlength{\parskip}{3.5pt}
\title{\vspace{-1.4em}\bfseries A Calibrated Spectral Survey of $\LU$ at
the Hou--Wang--Yang Profile\\[0.2em]\large\mdseries
No second unstable eigenvalue in the odd axisymmetric sector
\textemdash{} technical summary (companion to preprint v1.1)}
\author{P.~Salmond}
\date{\today}
\begin{document}
\maketitle
\vspace{-1.8em}

\begin{center}\itshape\small
A from-scratch, gate-validated realization of $\LU$ reproduces the
Hou--Wang--Yang eigenvalue $\lambda_1=0.113142$ to six significant
figures and exhibits no second unstable eigenvalue in the odd
axisymmetric sector at a calibrated sensitivity --- evidence for
instability order one in that sector.
\end{center}
\vspace{0.4em}

\paragraph{Question} Beyond your verified $\lambda_1 = 0.11314203\ldots$,
does $\LU$ (arXiv:2509.25116, your eq.\ for the linearized operator)
have further eigenvalues with $\operatorname{Re}\lambda > 0$ --- the
\emph{instability order}, which sets the dimension of the nonuniqueness
family? We survey the odd axisymmetric sector (the sector of
$\lambda_1$) in floating point, with every claim gated and the
instrument's sensitivity calibrated on $\lambda_1$ itself. Everything
below is evidence-grade, not certified.

\paragraph{Operator realization (trust anchors)} A pointwise Python
realization of $\LU$ in your $(\beta,\theta)$ coordinates, evaluated
against your published repository data (read at run time; not
redistributed): profile-equation residual $6.2\times 10^{-9}$, stored
eigenpair residual $1.7\times 10^{-5}$, archive md5-gated. Exact
symmetry eigenvalues as a side result: $\LU(\partial_z U) =
-\tfrac12\,\partial_z U$ and $\LU(U + \xi\!\cdot\!\nabla U) =
-(U + \xi\!\cdot\!\nabla U)$, so no equation symmetry generates an
unstable mode.

\paragraph{Instrument} The gradient-deflated residual landscape
\[
r(\lambda) \;=\; \min_{v \in \mathcal{V},\, q}
\frac{\|(\LU^{\mathrm{mom}} - \lambda) v - \nabla q\|_{L^2}}{\|v\|_{L^2}}
\;=\;\sqrt{\lambda_{\min}\!\bigl(\tilde A_d
 - \lambda \tilde G^{\!\top} - \bar\lambda \tilde G
 + |\lambda|^2 I\bigr)},
\]
the second-order-relative-spectrum / residual-pseudospectrum distance
function (the quantity underlying residual-DMD spectral validation;
Levitin--Shargorodsky, Trefethen--Embree, Colbrook--Townsend), here
realized for the divergence-free Navier--Stokes structure: $\mathcal{V}$
is an exactly divergence-free
(Stokes-stream) trial space, the scalar span absorbing $\nabla q$
replaces the Leray projection, and the pencil is exactly quadratic in
$\lambda$ because $\langle w, \nabla q\rangle = 0$ on our quadrature
domain (verified to $4\times 10^{-13}$). A genuine eigenvalue whose
eigenfunction the span represents produces a dip; the dip's attainable
depth is governed by the law below. (Motivation: in our first survey,
on the $\phi$-subspaces of your repository, a flagged second value
$+0.036$ passed residual, jackknife, and negative-control checks yet
was unmasked by this landscape as a pseudospectral shadow of
$\lambda_1$ --- documented in the preprint as a cautionary case study.)

\paragraph{Two numerical-analysis facts (known technique, silent traps)} Neither is new in isolation --- weighted quadrature for axis/pole singularities is standard since Bernardi--Dauge--Maday (1999), pressure elimination by divergence-free bases is textbook --- but both are invisible at the Rayleigh--Ritz pencil level and fatal to the operator-side Gram.
(i) \emph{Open-domain exactness}: for div-free decaying $w$, the
truncated-domain $\langle w, \nabla p\rangle$ equals the cut-edge flux
exactly (numerical closure $2.3\times 10^{-13}$); the standard
equatorial standoff $\theta \le \pi/2 - 0.04$ carries the
\emph{entire} pressure term ($p$ is odd about the equator;
$96\%$ of the operator output), while open-endpoint Gauss--Legendre on
$(0,\pi/2)^2$ kills all four fluxes analytically --- the dropped term
collapses $6.5\times 10^{4} \to 1.3\times 10^{-8}$.
(ii) \emph{Operator domain}: radial factors with $R(0) \ne 0$ (including
shell-Hermite $\psi_n$ inside the turning point) give
$\|\LU w\|^2 \sim \int \rho^{-2} d\rho = \infty$: the pencil is fine
but the operator-side Gram does not exist. The mask
$m(\rho) = \rho^2/(\rho^2 + 4)$ on all radial factors (admissible for
any $R$ in the Stokes-stream construction) cures it.

\paragraph{Trial space designed from the measured target} On the open
quarter domain the sector's Legendre families are orthogonal by
parity, so $v_1$'s angular spectrum is an exact Parseval decomposition:
the tail crosses $3\times 10^{-4}$ at $k = 102$. Per-$k$ least squares
under the divergence-free tie show a dual-width radial requirement
($\sigma = 2.5 \oplus 1.2$ shell-Hermite; single-width floor
$8.4\times 10^{-4} \to 7\times 10^{-5}$), and license a taper
(toroidal: 30 radial functions; poloidal $k \ge 40$: 46). Result:
$N = 5366$, $v_1$ fit error $2.557\times 10^{-4}$, within a
4\,GB/4-core budget.

\paragraph{Gate ladder (selection)}
\begin{center}\small
\begin{tabular}{lll}
\toprule
separable/heterogeneous assembly vs.\ direct operator columns &
$\le 5\times 10^{-15}$ & T4/T5/T6 \\
Ritz recovery of $\lambda_1$ (two regularization cuts) &
$|\mathrm{d}| = 3.0\times 10^{-4}$ & X1 \\
$U \equiv 0$ control: $\max\operatorname{Re} \le -\tfrac14$ &
$-0.61$ & X3 \\
pressure-deflation floor on the exact eigenpair &
$7.6\times 10^{-5}$ & calibration \\
deflated operator Gram bottom, cut-stability & 7 digits & PSD \\
\bottomrule
\end{tabular}
\end{center}

\paragraph{Resolving power and calibration} The landscape floor obeys
$r(\lambda_1) \lesssim \varepsilon_{\mathrm{pt}} + \mathcal{A}\,\delta$
with $\delta$ the span's best-approximation error of the eigenfunction
and $\mathcal{A}$ the operator amplification of the missing content.
Measured: $\delta = 4.4\times 10^{-3} \Rightarrow$ flat landscape at
$0.066$ (no valley); $\delta = 2.6\times 10^{-4} \Rightarrow$ valley of
depth $9.3\times 10^{-3}$. The known $\lambda_1$ thus calibrates both
the attainable depth and the location accuracy for any candidate.

\paragraph{Results} Real axis $[-0.10, 0.30]$: a single smooth basin;
fine-scan vertex at $\lambda = 0.113142$ ($\lambda_1$ to six
significant figures); depth $r = 9.3141\times 10^{-3}$, reproduced
\emph{pointwise} by an independent path (explicit arg-min field, one
direct operator application, grid-level deflation: $9.3141\times
10^{-3}$); raw values stable to $1.5\times 10^{-4}$ relative across
cuts, so $\gtrsim 1\%$ local dips are detectable. Complex window
$\operatorname{Re}\lambda \in [0,0.35]$, $\operatorname{Im}\lambda \in
[0.05,0.60]$: $r$ strictly monotone in $\operatorname{Im}\lambda$ on
every grid row; minimum $1.24\times 10^{-2}$ adjacent to the real-axis
valley; no off-axis local minimum. \emph{Conclusion (scoped): no
second unstable eigenvalue of $\LU$ in the odd axisymmetric sector at
the instrument's stated sensitivity; instability order one in this
sector.} Untouched: the even sector (surveyed only with the weaker
$\phi$-span), all $m \ge 1$ sectors, and rigorous certification.

\paragraph{What is and is not new here} The instrument is the residual-pseudospectrum / ResDMD validation principle applied to this operator, not a new method; the two facts above specialize standard axisymmetric-spectral practice. The elements we claim as new are the \emph{resolving-power law with its calibrate-on-$\lambda_1$ protocol} (turning a null survey into a quantified-sensitivity statement), the first spectral reconnaissance of $\LU$ beyond $\lambda_1$, and the documented false-positive case study.

\paragraph{What we would value your view on} (1) Whether this survey
and its negative are useful to your program and the community, and any
errors you see, particularly in our reading of your data conventions.
(2) Whether your certified finite-rank machinery could turn the
sector-scoped negative into a rigorous enclosure. (3) Whether the data
usage and attribution are as you would wish (your repository carries no
license file; we redistribute none of it). (4) If you judge the work
worth posting, we would be grateful for assistance with arXiv
submission (endorsement for math.AP).

\paragraph{Pointers} Preprint v1.1 (12 pp.) and the full pipeline
--- gates, forensics, reproduction ladder with expected numbers ---
at \url{https://github.com/psalmond/hwy-spectral-survey}.
\end{document}
